Architectures and Open pNets

### 2018-04-23 2018-Arch-pNets/arch-pNets.tex
Imported the example from the SMT paper Extended with data -- only in the figure Text still to update

@@

@@

%****************************************************************
\subsection{Graphical representation of composed systems}
\label{secn:connectors}

@@

\bibliographystyle{abbrv}
\bibliography{biblio.bib}

>>>>

\input{figures/macros.tex}         % Definitions of macros used in XFig figures

@@

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
This approach ensures that any property, shown to hold on a given BIP
model, will also hold by construction on the generated code.

In this context, the notion of \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, thereby potentially introducing additional
coordinator components and modifying the connectors that define
synchronisation patterns among ports of components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie when several architectures are applied, each enforcing a safety
property, the resulting system satisfies their conjunction.

The concepts and results of the paper are illustrated by a running
example extending the Failure Monitor architecture from the CubETH
nanosatellite on-board software case-study~\cite{CubETH-case-study}.

@@

%****************************************************************
\subsection{Running example: Failure Monitor architecture}
\label{secn:failure-monitor}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{BIPspec-ArchFailureTimerMax-v3}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{schema:ArchFailure:BIP}
\end{figure}

Figure~\ref{schema:ArchFailure:BIP} shows a refined version of the
Failure Monitor architecture used in~\cite{CubETH-case-study}.
Contrary to standard BIP models, architectures comprise one or several
\emph{operand} components, whereof only the set of \emph{ports} is
given.  Here, the operand component is $\NameOprnd$ and its interface
consists of the ports $\PortFinish$, $\PortResume$ and $\PortFail$.
The two \emph{coordinator} components\mdash $\NameCtrl$ and
$\NameTimer$\mdash are standard BIP components insofar as they also
have their \emph{behaviour} specified by finite automata extended with
local data variables.  Transitions of these automata are labelled with
the ports of the corresponding components, Boolean guards and update
functions on local variables.  For instance the loop transition
$\tstate{1} \goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{1}$
in the $\NameTimer$ component is labeled by the port $\PortTick$, it
can be fired only when the current value of the local variable $t$ is
greater than $0$.  Upon firing, this transition decrements the value
of $t$ by $1$.  When omitted, the default guard (\resp update
function) is the constant predicate $\true$ (\resp the $\mathit{skip}$
operator).  The constant $\mathrm{Max}$, in $\tstate{0}
\goesto{\PortStart, \TimerInit[]} \tstate{1}$, is a parameter of the
architecture.

Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of \emph{interactions}, based on the
attributes of the connected ports~\cite{BliSif08-acp-tc}, which may be
either \emph{trigger} (triangles in Fig.~\ref{schema:ArchFailure:BIP})
or \emph{synchron} (bullets in Fig.~\ref{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction may be executed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors,
\emph{involving at least one of the triggers}, can be executed.
%
For instance, the two ports $\NameTimer.\PortStart$ and
$\NameCtrl.\PortFail$ are always synchronised, since they belong to
the same binary sub-connector, where they are both synchrons.  In
particular, this means that whenever the transition $\cstate{0}
\goesto{\PortFail} \cstate{1}$ is fired, so is the transition
$\tstate{0} \goesto{\PortStart, \TimerInit[]} \tstate{1}$,
initialising the timer.  The binary connector $\NameTimer.\PortStart
\bullet\!\!\!-\!\!\!-\!\!\!\bullet \NameCtrl.\PortFail$ is a
sub-connector of a hierarchical connector, where the port
$\NameOprnd.\PortFail$ is a trigger.  Thus, the above interaction can
only happen together with $\NameOprnd.\PortFail$, forming a
ternary interaction.  On the contrary, being a trigger, the port
$\NameOprnd.\PortFail$ can fire alone, forming a singleton
interaction.  The composition semantics of BIP systems consists in
firing exactly one interaction, enabled through at least one of the
top-level connectors, at each execution round.

Finally, \emph{priorities}\mdash defined by a strict partial order on
the set of possible interactions\mdash narrow the choice among the
enabled interactions at any given round.  The default priority is the
so-called \emph{maximal progress}, whereby among any two interactions
$a \subset b$ (as sets of ports), $b$ has higher priority than $a$.
For example, the port $\NameOprnd.\PortFail$ will never fire alone in
a global state, where both $\NameTimer.\PortStart$ and
$\NameCtrl.\PortFail$ are enabled.
%
%% Adding priorities $\NameOprnd.\PortFinish < \NameCtrl.\PortReset$ and
%% $\NameOprnd.\PortFinish < \NameOprnd.\PortResume$, ensures that, after
%% a failure, the operation performed by $\NameOprnd$ cannot finish
%% unless it is explicitly resumed or the system is reset.

Application of the Failure Monitor architecture ensures that, whenever
a failure is registered in the operand component, the system will be
reset, unless a resumption is registered within $\mathrm{Max}$ time
units (more details in Sect.~\ref{section:resultOA}).

@@

\bibliographystyle{abbrv}
\bibliography{biblio}
